The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-07-18 paper.tex
Added SonRise section to Irish

Epenthesis does not break up onset sonorant-initial clusters.  This is likely due to the fact that
with a small number of lexical exceptions, Irish words are stress-initial \citep[26]{o.siadhail.1989}.\footnote{With the exception of Munster Irish, where stress is attracted to syllables containing
a long vowel or diphthong (Green 1996), %TODO
but where onset clusters still are not broken up by
epenthesis.  In this case, one would probably have to posit a positional faithfulness constraint
protecting the word-initial cluster.}  
Breaking up the onset cluster would result in an unstressable epenthetic vowel occupying the 
preferred location of stress in Irish.

@@

In codas and across syllable boundaries, epenthesis does not apply to homorganic clusters.  
I suggest that this is due to the action of the Obligatory Contour Principle (OCP).  
I adopt Walter 2007's %TODO
formulation of the OCP, which bans successive identical oral gestures.  A cluster such as [nt] would consist of a single oral constriction gesture at the alveolar ridge, and would not constitute a violation of the OCP. Splitting it by epenthesis to form [n\textipa{@}t] would require a repeated identical constriction gesture, and therefore be a violation of the OCP.

@@

\ex. \a. [\textipa{s\textsuperscript{j}e:mt}] \\
         {\it s\'eimt} `the act of playing music', dialectal variant of {\it seinm})
     \b. [\textipa{k2r\textsuperscript{j}@m\textsuperscript{j}k\textsuperscript{j}}] \\
         {\it cuirimc} `I observe, pay attention to' (Connaught dialect) \\
     (Words found in \citet{dinneen.2007}) %TODO

I conclude that in general, epenthesis does not occur in /NT/ sequences due to the large sonority angle between NT and NVT, and not to the non-existence of underlying heterorganic /NP/ clusters in the lexicon.

>>>>

Epenthesis does not break up onset sonorant-initial clusters.  This is likely due to the fact that
with a small number of lexical exceptions, Irish words are stress-initial \citep[26]{o.siadhail.1989}.\footnote{With the exception of Munster Irish, where stress is attracted to syllables containing
a long vowel or diphthong \citep{green.1996},
but where onset clusters still are not broken up by
epenthesis.  In this case, one would probably have to posit a positional faithfulness constraint
protecting the word-initial cluster.}  
Breaking up the onset cluster would result in an unstressable epenthetic vowel occupying the 
preferred location of stress in Irish.

@@

In codas and across syllable boundaries, epenthesis does not apply to homorganic clusters.  
I suggest that this is due to the action of the Obligatory Contour Principle (OCP).  
I adopt \citet{walters.2007}'s
formulation of the OCP, which bans successive identical oral gestures.  A cluster such as [nt] would consist of a single oral constriction gesture at the alveolar ridge, and would not constitute a violation of the OCP. Splitting it by epenthesis to form [n\textipa{@}t] would require a repeated identical constriction gesture, and therefore be a violation of the OCP.

@@

\ex. \a. [\textipa{s\textsuperscript{j}e:mt}] \\
         {\it s\'eimt} `the act of playing music', dialectal variant of {\it seinm})
     \b. [\textipa{k2r\textsuperscript{j}@m\textsuperscript{j}k\textsuperscript{j}}] \\
         {\it cuirimc} `I observe, pay attention to' (Connaught dialect) \\
     (Words found in \citet{dineen.2007})

I conclude that in general, epenthesis does not occur in /NT/ sequences due to the large {\sc Sonority Angle} between NT and NVT, and not to the non-existence of underlying heterorganic /NP/ clusters in the lexicon. The epenthetic facts of Irish thus support the ranking of NT and RT clusters at the bottom of the hierarchy of susceptibility to epenthesis, which is consistent with the {\sc Sonority Angle} metric.

\subsection{Comparison to other approaches}

Using the same revised sonority scale, the {\sc Sonority Rise} calculations are as follows for the relevant clusters:

\ex. {\sc Sonority Rises} of sonorant-initial clusters using revised sonority scale

 \begin{center}
  \begin{tikzpicture}[shorten >=1pt,->,scale=0.5]
     \tikzstyle{line} = [draw]%, -latex']

        \node (NT) at (1.66666666666667 * 15, 0) {NT}; 
        \node (ND) at (1.5 * 15, 0) {ND}; 
        \node (NF) at (1.33333333333333 * 15, 0) {NF};
        \node (NN) at (1 * 15, 0) {NN};
        \node (NR) at (0.666666666666667 * 15, 0) {NR};

        \node (RT) at (2.5 * 15, 0) {RT};
        \node (RD) at (2.25 * 15, 0) {RD}; 
        \node (RF) at (2 * 15, 0) {RF}; 
        \node (RN) at (1.5 * 15, 0.7) {RN};
        \node (RR) at (1 * 15, 0.7) {RR};

    \node (axisstart) at (0.5 * 15,-1) {};
    \node (axisend)   at (2.6 * 15,-1) {};
    \draw (axisstart) -- (axisend);
   \node (xaxislabel) at (2.6 * 15,-1.5) {\textsc{Sonority Rise}};

    \end{tikzpicture} % TODO: draw arrow and mark 1.0, 1.5, 2.0, 2.5
 \end{center}

Note that the clusters RF and RD fall between RT and NT on this scale.  Therefore, we cannot clearly separate the clusters that undergo epenthesis
from the ones that do not if we adopt {\sc Sonority Rise} as the relevant metric.
